\customchapter{Post Install Configurations}{System identity, licensing, and workstation preparation}

In this chapter we will discuss the post-installation configurations required for the deployed workstations. This includes setting up the appropriate hostname and username nomenclature aligned with the license provided by C2S, preparing the system identity used by the Cadence environment, and installing the relevant software packages required for the Cadence Suite.

\section{Username/Hostname Configuration} 

After the base RHEL installation is complete, the first administrative task is to standardize the
identity of each workstation. In an EDA laboratory environment, usernames and hostnames should not
be chosen casually. They are used by the operating system, network services, license checks, remote
administration scripts, shared storage mounts, and Cadence startup environments. If different
machines follow different naming patterns, the environment becomes difficult to audit and errors in
license or path configuration become much harder to diagnose.\\

According to the license our university has received, the workstation identity should follow the
following naming convention:
\[
	\text{user0\_X} \; @ \; \text{ECE-X.gsv.ac.in}
\]

\noindent Here, \texttt{user0X} represents the local Linux username assigned to the workstation user, while
\texttt{ECE-X.gsv.ac.in} represents the fully qualified hostname assigned to that workstation on the
department network. The value of \texttt{X} should be replaced by the workstation number or seat number
allocated by the laboratory administrator. For example, workstation 3 may use the username
\texttt{user003} and the hostname \texttt{ECE-3.gsv.ac.in}. The important point is that the same numbering
scheme must be used consistently across the user account, hostname records, license configuration,
and any administrative documentation.\\

This convention gives every machine a predictable identity. A system administrator can immediately
identify which physical workstation corresponds to which Linux account and hostname, and the Cadence
license environment can be configured against a stable naming scheme. Before installing or launching
Cadence tools, confirm that the selected username and hostname match the C2S license documentation
and the laboratory workstation inventory. If a machine is renamed later, related configuration files,
license references, and shell startup scripts may also need to be updated.

\begin{tcolorbox}
At the time of writing this document, there are officially upto 100 registered workstations for GSV students under the C2S License. The first 70 systems are reserved for the ECE department, while the remaining 30 are reserved for EE department.
\end{tcolorbox}

This very first priority is ensuring the groupname and username/hostnames are as per the respective workstations. File ownership depends heavily on the groupname and GID (group ID). In Linux, the default file permissions and ownership records are verified using GID and UID (user ID), hence ensuring they agree, resolves problems down the line.\\

On a fresh install, usually the default group ID is set to \textbf{1000} for the default user. In our case it is \textbf{user003}. You can verify this with the following commands:

\begin{codeblock}{terminal}
id user003  && whoami && hostname
\end{codeblock}


The output should display the expected username, User ID (UID), group name, Group ID (GID), and hostname associated with the workstation. These values should be verified before proceeding with the Cadence installation, as several configuration files and environment settings may reference the user account, host identity, or group ownership. If any of these values are incorrect, they should be corrected prior to continuing with the software installation process.\\

\noindent Once the workstation identity has been verified, the next step is to refresh the package metadata and update the operating system packages. On RHEL 9.8, this is typically performed using the \texttt{dnf} package manager. The following command updates all installed packages to their latest available versions:

\begin{codeblock}{terminal}
sudo dnf update -y
\end{codeblock}

After the update process is complete, the workstation should be rebooted if any kernel, driver, or system library updates have been installed. This ensures that the running operating system matches the updated package state before any Cadence-specific dependencies or environment configurations are applied. \\

The primary configuration tasks at this stage involve setting up the required environment variables, configuring executable search paths, and ensuring that the appropriate file permissions are assigned to the intended user account. Since multiple workstations must be configured with identical software environments, a system imaging and cloning approach is employed using Clonezilla. \\

After a workstation image has been cloned using Clonezilla, the target machine inherits several machine-specific attributes from the source workstation, including the username, hostname, home directory structure, and file ownership information. These parameters must be customized before the workstation can be deployed for regular use.\\

Traditionally, this process would require the administrator to manually perform the following tasks:

\begin{itemize}
	\item Renaming the user account
	\item Renaming the primary group
	\item Updating the user's display name
	\item Updating the hostname
	\item Updating /etc/hostname
	\item Updating /etc/hosts
	\item Correcting ownership of user directories
\end{itemize}

Performing these operations manually on every cloned workstation is repetitive and increases the likelihood of configuration errors. To streamline the deployment process and ensure consistency across all laboratory systems, these tasks have been automated through a dedicated C-shell (\texttt{csh}) configuration script. The script prompts the administrator for the required workstation-specific parameters and automatically applies the necessary user, group, hostname, and ownership modifications, thereby reducing deployment time and minimizing the risk of human error. \\

The full script is present in Appendix~\ref{app:c2s}


\section{Installing and Configuring Cadence}

\subsection{Cadence Directory Structure}
Cadence expects everything to be present in the \inlinecode{/home/installs} directory, so we need to make that: 


\begin{codeblock}{terminal}
sudo mkdir -p /home/installs
\end{codeblock}

This is the place where we install all the tools in the Cadence Suite and unzip them. 

For a laboratory deployment, the \inlinecode{/home/installs} directory should be treated as a
centralized local software repository rather than as an arbitrary extraction location. All Cadence
releases, auxiliary setup files, and vendor-provided directory trees should be placed under this
common root so that every workstation follows the same filesystem layout. This standardization is
important for system imaging, recovery, and later maintenance. If one workstation stores the Cadence
tree in a different location, the shell startup file, environment variables, license diagnostics, and
teaching scripts may need to be modified only for that machine, defeating the purpose of a common
laboratory image.\\

\noindent The Cadence installation archives used for the C2S laboratory setup are obtained from the
repository provided for the deployment. Administrators should download the packages from the
following source and preserve the original archive structure supplied there:

\begin{center}
\small\url{https://entupletech-my.sharepoint.com/personal/cadence_support_entuple_com/_layouts/15/onedrive.aspx?id=%2Fpersonal%2Fcadence%5Fsupport%5Fentuple%5Fcom%2FDocuments%2FDATA%2FCadence%5FRHEL%5F8%5Fversion%2FIndividual%5Flinks&ga=1}
\end{center}

\noindent In normal workstation preparation, the archive is first downloaded into the user's
\inlinecode{~/Downloads} directory. This keeps the downloaded source material separate from the
final installation location and allows the administrator to inspect the archive before extraction.
After verification, the archive must be extracted directly into \inlinecode{/home/installs/}. No
additional wrapper directory should be created during extraction. The contents of the archive should
land directly inside \inlinecode{/home/installs/} so that the Cadence directory hierarchy expected
by the later environment-variable, C-shell initialization, and license setup steps is preserved.\\

\noindent The exact archive name may differ depending on the release downloaded from the repository.
The following example therefore uses a clearly marked placeholder for the archive filename:

\begin{codeblock}{terminal}
cd ~/Downloads
ls -lh

# Replace this with the actual archive downloaded from the repository.
ls -lh <CADENCE_ARCHIVE_NAME>

sudo tar -xf <CADENCE_ARCHIVE_NAME> -C /home/installs/

ls -lh /home/installs
find /home/installs -maxdepth 2 -type d | sort
\end{codeblock}

\noindent If the downloaded file is provided in a different compressed format, the extraction command
should be adjusted accordingly while keeping the destination unchanged. For example, a
\inlinecode{.zip} archive should still be extracted with \inlinecode{/home/installs/} as the target
directory. The administrative rule remains the same: do not create an extra directory layer between
\inlinecode{/home/installs/} and the Cadence tree supplied by the archive.\\

\noindent Preserving the original hierarchy is critical because subsequent configuration steps refer
to predictable paths below \inlinecode{/home/installs}. The \inlinecode{.cshrc} file, Cadence
environment variables, executable search path, runtime library path, and license diagnostic commands
all assume that the installation tree has not been rearranged manually. A small change made during
extraction can later appear as a shell configuration failure or a license startup problem, even
though the root cause is simply that the expected directory structure was not preserved. For this
reason, the extraction layout should be verified immediately after unpacking the archive and before
any environment variables are written.

\subsection{Environment Variables}

Before launching any Cadence tool, the user's shell environment must point to the correct
installation directories and license server. If a working \inlinecode{.cshrc} file is already present
and contains the required paths, it does not need to be recreated. The important requirement is that
the active shell can locate the Cadence executables, libraries, setup files, and license
configuration.\\

\noindent The following values should be checked for every workstation:

\begin{itemize}
	\item The Cadence installation root under \inlinecode{/home/installs}
	\item The executable paths for Cadence tools, usually the \inlinecode{bin} and \inlinecode{tools/bin}
	      directories inside the Cadence installation
	\item The runtime library path required by the installed Cadence release
	\item The license variables, such as \inlinecode{CDS_LIC_FILE} and \inlinecode{LM_LICENSE_FILE},
	      pointing to the correct license server and port
	\item Any release-specific setup paths required by the installed Cadence package
\end{itemize}

\noindent The complete C-shell startup configuration is provided in the appendix:
\ref{app:csh}. After confirming or updating the startup file, log
out and log back in, or source the file manually, and verify the active environment using:

\begin{codeblock}{terminal}
echo $CDS_INST_DIR
echo $CDS_LIC_FILE
which genus
\end{codeblock}

If \inlinecode{which genus} does not return a path under the Cadence installation directory, the
executable path has not been configured correctly. The environment variables should be corrected
before continuing with license checks or tool validation.\\

\subsection{C-Shell Configuration}

The C-shell startup file, usually stored as \inlinecode{~/.cshrc}, is one of the most important
configuration files in a Cadence workstation deployment. It is read whenever an interactive C-shell
session is started, and it defines the environment that the user receives before any EDA tool is
launched. In a single-user Linux installation this file may only contain a few aliases or display
preferences, but in an EDA laboratory it becomes part of the workstation infrastructure. It controls
which tool release is visible, which license server is contacted, which library paths are searched,
and whether the shell behaves consistently across cloned systems.\\

\noindent Cadence installations have traditionally relied on C-shell based initialization because many
vendor-provided setup scripts, examples, and legacy flows assume \inlinecode{csh} or
\inlinecode{tcsh} syntax. Even when the operating system default shell is Bash, laboratory users may
still be instructed to start Cadence tools from a C-shell session so that the same environment logic
is applied on every workstation. This reduces variation between machines and prevents accidental
differences between users who may otherwise maintain separate Bash, Zsh, or graphical terminal
profiles. For a C2S laboratory, where the same toolchain must be reproduced across many student
systems, this consistency is more important than personal shell preference.\\

\noindent The \inlinecode{.cshrc} file should therefore be treated as a controlled configuration file.
Administrators should avoid scattering Cadence variables across multiple unrelated startup files,
because such a layout makes troubleshooting difficult after cloning or system repair. A maintainable
configuration should clearly define the installation root, append the required executable directories
to the shell path, define any required runtime library locations, and set the licensing variables
used by Cadence and FlexLM. If several Cadence releases are installed under \inlinecode{/home/installs},
the selected release should be documented in the file so that future administrators can identify
which version is active without inspecting every directory manually.\\

\noindent A minimal administrative check is to confirm that the user account is actually reading the
expected C-shell startup file. This should be tested from the same account that will run the tools,
not only from the root account, because file ownership, home directory changes, and cloned user
profiles can easily produce different results for different users. The following commands may be
used after logging in as the workstation user:

\begin{codeblock}{terminal}
echo $shell
ls -l ~/.cshrc
source ~/.cshrc
rehash
\end{codeblock}

\noindent The \inlinecode{source} command reloads the startup file in the current shell, while
\inlinecode{rehash} refreshes the C-shell command lookup table after new executable directories have
been added to the path. This is useful during installation because the administrator can edit the
startup file and test it immediately without logging out after every change. However, final
validation should still be performed in a new login session, because that confirms that the settings
are loaded automatically and do not depend on manual commands typed during installation.\\

\noindent The C-shell configuration should not contain workstation-specific values unless those values
are intentionally different for each machine. The tool installation path and license configuration
should normally be identical across all cloned workstations, while the username and hostname are
handled separately during the system identity configuration described earlier in this chapter. This
separation improves maintainability: a cloned machine can receive its own hostname and user identity
without requiring a different Cadence startup file for every seat in the laboratory.

\subsection{License Configuration}

Cadence tools are protected by network licensing, commonly implemented through the FlexLM licensing
system. In this model, the workstation does not independently decide whether a tool may run. Instead,
when a Cadence executable starts, it contacts the configured license server and requests the required
license feature. The server checks whether the feature exists, whether the request is allowed, and
whether enough license tokens are available. Only after this exchange succeeds does the tool proceed
to the normal graphical or command-line startup sequence.\\

\noindent License configuration is therefore a core infrastructure dependency, not a cosmetic
environment setting. A workstation may have a complete Cadence installation and a correct executable
path, but the tools will still fail to launch if the license variable is missing, the server hostname
cannot be resolved, the port is incorrect, or the network path to the server is unavailable. In a
teaching laboratory this can affect many users at the same time, so license settings should be kept
centralized, documented, and reproducible. The same license configuration should be applied to every
workstation unless the C2S license documentation explicitly requires separate values for separate
systems.\\

\noindent The two variables most commonly involved are \inlinecode{CDS_LIC_FILE} and
\inlinecode{LM_LICENSE_FILE}. The Cadence-specific variable points Cadence tools to the license
source, while \inlinecode{LM_LICENSE_FILE} is understood by FlexLM-based utilities and may also be
used by other licensed software. The value normally follows the form
\inlinecode{PORT@LICENSE_SERVER}, where \inlinecode{PORT} is the FlexLM server port and
\inlinecode{LICENSE_SERVER} is the hostname or IP address of the license server. The actual value
must be taken from the C2S license documentation or from the laboratory license administrator. If
the value is not yet known during documentation preparation, it should be left as a clearly marked
placeholder rather than replaced by an assumed address:

\begin{codeblock}{csh}
setenv CDS_LIC_FILE <PORT>@<LICENSE_SERVER_HOSTNAME>
setenv LM_LICENSE_FILE $CDS_LIC_FILE
\end{codeblock}

\noindent Hostname resolution should be checked before blaming the Cadence installation itself. If
the license variable uses a hostname, that hostname must resolve through DNS or through a controlled
entry in \inlinecode{/etc/hosts}. If the laboratory network is segmented, administrators should also
confirm that the workstation can reach the license server from the student VLAN or laboratory subnet.
Using an IP address can avoid name resolution problems, but it also makes future server migration
less transparent. For maintainability, a stable hostname is usually preferable when the network
infrastructure supports it correctly.\\

\noindent Licensing failures commonly occur because the startup file was not sourced, the wrong shell
was used, the license server name was mistyped, the FlexLM port was changed, or a firewall blocked
the connection. Another frequent cause is a mismatch between the installed Cadence release and the
features available in the license file. For this reason, license validation should be performed after
every installation, after every workstation clone, and after any license server maintenance window.
The following checks are useful when the FlexLM utility is available in the active environment:

\begin{codeblock}{terminal}
echo $CDS_LIC_FILE
echo $LM_LICENSE_FILE
lmutil lmstat -a -c $CDS_LIC_FILE
\end{codeblock}

\noindent If \inlinecode{lmutil} is not available directly in the shell path, the administrator should
use the copy provided by the installed Cadence release, or replace the command with the site-approved
license diagnostic utility. The important point is not the exact utility name, but the validation
principle: the workstation must be able to contact the same license source that Cadence will use at
runtime.

\subsection{Verifying Installation}

Verification should be performed systematically before a workstation is released for student use.
The goal is not merely to prove that one command opens once, but to confirm that the deployed machine
matches the intended laboratory baseline. This includes the operating system identity, the Cadence
installation directory, the C-shell startup configuration, the executable search path, the runtime
library environment, and access to the license server. A structured validation step is especially
important after Clonezilla deployment, because cloned machines can inherit paths and user
configuration from the source image while still requiring workstation-specific identity changes.\\

\noindent The first validation step is to check that the environment visible to the user is the same
environment described in the configuration files. This must be done from a fresh terminal session as
the intended user account. Administrators should avoid relying only on commands run immediately
after manual sourcing, because that can hide startup errors that will reappear when students log in
normally. The following commands provide a basic view of the active shell and Cadence environment:

\begin{codeblock}{terminal}
whoami
hostname
echo $shell
echo $CDS_INST_DIR
echo $CDS_LIC_FILE
echo $LM_LICENSE_FILE
\end{codeblock}

\noindent The values should match the workstation identity convention and the Cadence configuration
approved for the laboratory. Empty output for a required variable usually indicates that the startup
file was not read, the user is not running the expected shell, or the variable was defined in a file
that is not part of the current login sequence. Incorrect output is more serious than empty output,
because it may point the user to an outdated release or to an obsolete license server without making
the failure obvious at first.\\

\noindent The next step is to confirm executable visibility. Cadence tools are large application
suites, and their front-end commands must be reachable through the C-shell path. If the shell cannot
find the executable, the installation may still exist on disk, but the user environment is not ready.
The administrator should check both command discovery and version reporting where the tool supports
it:

\begin{codeblock}{terminal}
which genus
which xrun
which innovus

genus -version
xrun -version
innovus -version
\end{codeblock}

\noindent Not every laboratory installation will include every Cadence product. A missing command is
therefore not automatically an error if that product was not installed or was not included in the
C2S entitlement. However, every tool expected for the course flow should resolve to a path under the
approved Cadence installation directory, and its version should correspond to the documented release.
If different machines report different versions, the clone image or startup file should be reviewed
before the laboratory is used for instruction.\\

\noindent License access should be validated after executable visibility, because a tool may be
installed correctly but still fail during feature checkout. If FlexLM diagnostics are available, the
license server can be queried before opening a full graphical application:

\begin{codeblock}{terminal}
lmutil lmstat -a -c $CDS_LIC_FILE
\end{codeblock}

\noindent Finally, the administrator should perform a practical launch test using at least one
graphical tool and one command-line or batch-capable tool from the expected design flow. This catches
problems that simple path checks may miss, such as missing shared libraries, display forwarding
issues, incompatible system packages, or permission problems in the user's working directory. The
workstation should only be considered ready when these checks succeed from the normal student
account without requiring root privileges.

\subsection{Launching Cadence Tools}

Cadence tools should be launched from a prepared user environment, not from an arbitrary terminal
state. In normal laboratory use this means that the user logs in to the workstation, opens a
C-shell-compatible terminal session, and starts the required tool from a project working directory.
The working directory matters because many Cadence applications create local run files, logs, lock
files, simulation outputs, or design database directories during startup. Launching tools from a
temporary or root-owned directory can create permission problems that are difficult for students to
diagnose later.\\

\noindent During the first execution, the tool performs several checks before the user reaches the
main interface. It reads the shell environment, locates the installation tree, loads shared
libraries, checks release-specific setup files, contacts the license server, and initializes user
configuration data. A delay during the first launch is therefore not unusual, especially on a newly
cloned workstation or after a user profile has been created. However, repeated failures at this stage
usually indicate an infrastructure problem: an incorrect path, a missing library, an unavailable
license feature, or an environment mismatch between the selected release and the active shell.\\

\noindent A typical launch test for the synthesis environment is performed with
\inlinecode{genus}. The administrator should run it from a writable directory owned by the user,
observe whether the application starts without license errors, and then close it normally after the
startup sequence is complete:

\begin{codeblock}{terminal}
mkdir -p ~/cadence_test
cd ~/cadence_test
genus &
\end{codeblock}

\noindent Successful startup confirms more than the presence of the \inlinecode{genus} executable.
It confirms that the shell path is active, the required runtime libraries can be loaded, the license
server is reachable, and the release selected in the shell matches the installation tree on disk. If
the tool reports license warnings or missing setup files, the workstation should not be handed over
as ready until those messages are understood. Warning messages that appear harmless during
installation often become recurring support issues once a full class begins using the machines.\\

\noindent Digital design and implementation tools may also be checked from the command line. Version
queries are useful for confirming the active release without starting a full project, while a later
course-specific flow can be used for deeper validation. The following commands provide a lightweight
launch-level check for commonly used Cadence tools, subject to the products actually installed under
the C2S deployment:

\begin{codeblock}{terminal}
xrun -version
genus -version
innovus -version
\end{codeblock}

\noindent A successful launch or version response should be recorded as part of the workstation
handover checklist. In a laboratory setting, this record helps distinguish between installation
problems and later user-level project errors. Once the major tools start from the normal user
account, the expected license source is reachable, and the reported versions match the documented
release, the Cadence environment can be considered ready for instructional use.
