\newpage
\begin{center}
{\headingfont\LARGE\bfseries Abstract} \\
\end{center}
\addcontentsline{toc}{section}{Abstract}


\textit{this requires further enhancement} \\

  This manual covers the procedures required to deploy, provision, operate, maintain, and recover a RHEL-based semiconductor CAD workstation environment. It is intended to serve as an operational reference for administrators responsible for preparing workstations, installing EDA tools, configuring user access, and maintaining a reliable design environment over time. \\

  \noindent Specifically, this manual covers the following areas:
  \begin{itemize}
    \item Operating system deployment for RHEL workstations, including installation planning and baseline system configuration.
    \item Workstation provisioning, including package installation, network readiness, storage preparation, and post-installation setup.
    \item EDA installation workflows for supported semiconductor design tools and their required dependencies.
    \item Licensing configuration, including license server connectivity, environment variables, and validation steps.
    \item Shell environments for CAD users, including startup files, module or tool setup scripts, and reproducible environment initialization.
    \item User management tasks, including account creation, group membership, access control, and basic permission practices.
    \item Troubleshooting procedures for common operating system, EDA tool, license, shell, and user-access issues.
    \item Recovery procedures for failed installations, broken environments, workstation rebuilds, and service restoration.
    \item Maintenance workflows, including updates, backups, configuration audits, cleanup tasks, and periodic validation of the CAD environment.
  \end{itemize}

  \noindent This manual does not cover semiconductor design methodology, RTL design, verification strategy, physical design theory, circuit design, device modeling, or EDA tool usage from a design-engineering perspective. It also does not define institutional IT policy, network architecture outside the workstation environment, procurement procedures, vendor contract negotiation, or license entitlement management beyond the technical configuration needed for workstation operation.\\

\newpage
